How to make an abridged life table and calculate life expectancy

### 2016-09-16 working_teaching_examples/how_to_abridged_life_table.Rmd
made folder an rstudio project and changed life table code

```{r}
#dataset needs to be in order of age
life.table <- function(data, calendar.years, age.groups, num.ages.in.group, death.counts, population.counts){

  data["R_x"] <- data[death.counts]/data[population.counts] #mortality rates
  data["a_x"] <- 0.5 #average proportion of interval lived
  data[1, "a_x"] <- 0.09
  data["q_x"] <- data[num.ages.in.group]*data["R_x"]/(1 + (1 - data["a_x"])*data[num.ages.in.group]*data["R_x"]) #probability of dying
  data[dim(data)[1], "q_x"] <- 1 #100% die in last age group
  data["p_x"] <- 1 - data["q_x"] #probability of survival

  data["l_x"] <- NA #l_x: is the number of individuals alive at the start of the interval
  data[1, "l_x"] <- 100000

  for(i in 2:dim(data)[1]) {
    data[i, "l_x"] <- data[i - 1, "l_x"]*data[i-1, "p_x"]
  }

  data["d_x"] <- data["l_x"]*data["q_x"] #number of deaths
  data["L_x"] <- (data["l_x"] - data["d_x"])*data["n_x"] + data["d_x"]*data["a_x"]*data["n_x"] #person time in the age bracket
  data[dim(data)[1], "L_x"] <- data[dim(data)[1], "l_x"]/data[dim(data)[1], "R_x"] #different for last age bracket

  #now to calculate the total person time lived at or after age x, T_x:
  data["cumsum"] <- cumsum(data["L_x"])
  data["l.cumsum"] <- c(0, data[1:(dim(data)[1] - 1), "cumsum"])
  data["T_x"] <- sum(data["L_x"]) - data["l.cumsum"]
  data["e_x"] <- data["T_x"]/data["l_x"]

  return(data)

}

life.table.cm <- life.table(data = canada.men, age.groups = "ages", num.ages.in.group = "n_x", death.counts = "obs.deaths", population.counts = "obs.popn")

>>>>

```{r calculate_life_expectancy}
life.table.cm <- life.table(data = canada.men, age.groups = "ages", num.ages.in.group = "n_x", death.counts = "obs.deaths", population.counts = "obs.popn")
